\begin{abstract}
When the environment drifts, safe RL often fails before average reward collapses: the policy starts acting twitchy while its value estimates are still noisy, which is an awkward failure mode if you care about ``bad days'' (attacks, bursts, latency spikes) and not only about the mean curve.

We built a thin layer on top of PPO that runs at decision time. A calibrated uncertainty score picks between the learned action and a fixed heuristic; a separate shield clamps or rejects actions before they hit the plant. Nothing here is about beating every baseline on return---the point is architectural. Gating plus shielding shifts where the safety--stability trade-off sits compared with ``policy only,'' and we use a small, classically simulated VQC-style head so that routing signal does not cost an ensemble-sized footprint.

Most of the evidence is routine constrained-gym work. Each standard Safety Gym task in the protocol uses thirty seeds (task IDs appear in the setup section). The custom governance simulators are there as stress tests, again with matched seeds (fifty for the controller comparison block). We also print an embarrassing line on purpose: in one governance ablation the full stack loses to random on scalar reward. That stays in because reward alone is a weak certificate here---we care about stable, constraint-respecting behavior under shift and tight model budgets, not a universal return trophy.

Repository artifacts and scripts are available upon request.
\end{abstract}
