\section{Experimental Setup}

\paragraph{Protocol.}
We split evaluation into two buckets on purpose. The first bucket is vanilla constrained RL: SafetyPointGoal1-v0 and SafetyHalfCheetahVelocity-v1 with thirty seeds, sharing the same gating and shield code paths, compared against CPO, P3O, and PPOLag under OmniSafe. The second bucket is meaner: a scaled governance simulator (twelve states, three actions, two-hundred steps, partial observability, injected telemetry faults) and a DeFi-style simulator (eight states, two actions, one-fifty steps, MEV and liquidity stress baked in).

Publication tables use fixed seed sets so paired comparisons stay honest under noise. Heavy PPO-family and ablation sweeps use sixteen paired seeds; the official constrained-baseline block uses fifty; uncertainty and efficiency comparisons use thirty so the headline cells are not dominated by variance.

\paragraph{Governance comparison method set.}
The governance comparison table lines up six controllers on the same protocol in both environments: Random, Heuristic, CPO-style, P3O-style, PPO-Lagrangian-style, and Robust QGate+Shield.

\paragraph{Environment realism extension.}
Both simulators accept optional CSV-driven traces (demand, volatility, MEV pressure, oracle noise). That lets us replay coarse historical shapes instead of only drawing synthetic drift. The trace-backed run in the paper uses Ethereum daily gas-price history from Etherscan to schedule intensity.

\paragraph{Primary Metrics.}
Reward mean and standard deviation across seeds, violation counts, target MAE, how often fallback fires, and a per-step latency / forward-FLOPs proxy.

Fallback rate counts any step where execution came from the deterministic controller, either because the gate routed away from the policy or because the shield rejected the policy move and substituted the heuristic.

\paragraph{Hypothesis Testing.}
PPO-family comparisons use permutation tests and effect sizes where we care about pairwise structure; constrained baseline suites compare final return, final cost, and episode length directly. Bootstrap 95\% intervals appear where intervals read more clearly than point estimates alone.

\paragraph{High-rigor constrained-baseline protocol.}
Fifty seeds, frozen environment knobs, and metrics declared up front (reward, safety cost, violations in hard regimes) keep the baseline ordering from jittering with a single lucky draw.

\paragraph{Fair-capacity quantum-inspired efficiency protocol.}
Matched-parameter controls (VQC-28 vs MLP-28) sit next to larger classical nets so we can separate ``different head'' from ``more parameters.'' We report both raw and parameter-normalized views.

\paragraph{Primary standard-benchmark protocol.}
The external-validity anchor is still the Safety Gym pair above under shared gating and shielding, three thousand steps per run on this host. Governance runs are stress tests for regime behavior that is hard to stress in the standard suites alone.

\paragraph{Robust PPO settings.}
Robust PPO keeps risk-aware shaping, quantum-uncertainty gating, and the contract-style shield. Uncertainty uses $K=10$ perturbations with Gaussian noise scale $0.01$. $\tau_q$ and $\lambda_{\mathrm{risk}}$ were tuned where switching is aggressive.

\paragraph{Robust ablation variants.}
Seven configurations isolate pieces: Full (gate+shield+robust policy), No quantum gate, No shield, No fallback, Classical PPO, Heuristic, Random.
