\section{Reproducibility Checklist}

Code for the experiments and for regenerating tables lives in the repo. Seeds are fixed for the publication-scale runs so you can diff two branches without rerolling noise. JSON and markdown artifacts are emitted straight from scripts; LaTeX tables are generated from those files rather than hand-typed. CI runs lint, types, tests, and a few health checks so obvious foot-guns fail early.

The long path from raw logs to camera-ready tables looks like this:
\begingroup
\scriptsize
\begin{verbatim}
python experiments/run_publication_suite.py
python experiments/run_detailed_suite.py
python experiments/run_adversarial_suite.py
python experiments/run_complexity_\
regime_suite.py
python scripts/analyze_research_results.py
python scripts/generate_benchmark_report.py
python scripts/generate_paper_tables.py
python scripts/generate_appendix_tables.py
python scripts/generate_paper_figures.py
python scripts/generate_adversarial_\
artifacts.py
python scripts/generate_adversarial_\
stress_sweep.py
python scripts/generate_complexity_\
artifacts.py
\end{verbatim}
\endgroup

\paragraph{Provenance policy.}
We tried to avoid ``magic numbers'' in \texttt{.tex}: if a digit appears in the main text, it should trace back to an artifact file. That habit hurts when you refactor code, but it pays back the first time a reviewer asks for a regeneration.

\paragraph{One-command replication.}
For day-to-day checks there is a shorter entry point:
\begingroup
\scriptsize
\begin{verbatim}
make reproduce
\end{verbatim}
\endgroup
That target runs the subset of generators behind the main tables and figures (constrained transfer, uncertainty, efficiency, adversarial perturbation, qubit scaling). The longer script list above is still the full kitchen-sink pipeline if you want every auxiliary plot.

\paragraph{Checklist links.}
Venue-style checklists are checked in under \texttt{docs/} if you need wording for a particular submission:
\begin{itemize}
    \item \texttt{docs/NEURIPS\_SUBMISSION\_CHECKLIST.md}
    \item \texttt{docs/ICLR\_SUBMISSION\_CHECKLIST.md}
    \item \texttt{docs/IEEE\_SUBMISSION\_CHECKLIST.md}
\end{itemize}
Code and artifacts are available upon request.

\paragraph{Environment notes.}
Python with pinned seeds and deterministic script order is the reference stack. Bit-identical floats across machines are not a promise we make; protocol-level reproducibility \emph{is}---commands, outputs, and validation hooks are documented.

The checklist framing here matches how NIST and related bodies talk about AI and cyber risk management in 2024--2025 guidance \cite{nistairmf2023,nistcsf2024}.
